The tracing of trade: exploring the patterns of olive oil production and distribution from Roman Baetica

### 2020-11-05 Paper/jaccardbaetica.tex
finished and submit

\address[ceipacadress]{Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, University of Barcelona, Barcelona, Spain}
\address[edadress]{School of History, Classic \& Archaeology, Room OOM.33, William Robertson Wing, Old Medical School, Teviot Place, University of Edinburgh, UK}
\address[ubadress]{DIDPATRI, Passeig de la Vall d'Hebrón, 171, Barcelona, Spain}

@@

To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} (Inferior and Superior) and \textit{Britannia} as consumption centres. They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres. Here, we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance. The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes. Finally, results pretend to highlight the organisation of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.

@@

This research was partly funded by the European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros, Ignacio Morer and Víctor M. Lozano for their useful suggestions.  
Data have been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}. Maps were performed using QGIS software v.3.14 `Pi'. Datasets and source code are freely available under Open licenses from https://github.com/Mcotsar/Baetica\_stamps.

>>>>

\address[ceipacadress]{Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, Universitat de Barcelona, Barcelona, Spain}
\address[edadress]{School of History, Classic \& Archaeology, Room OOM.33, William Robertson Wing, Old Medical School, Teviot Place, University of Edinburgh, UK}
\address[ubadress]{DIDPATRI, Universitat de Barcelona, Passeig de la Vall d'Hebrón, 171, Barcelona, Spain}

@@

To achieve this goal, we analyse a set of stamps from two types of centres (production and consumption): 1) production centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} (Inferior and Superior) and \textit{Britannia} as consumption centres. They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres. Here, we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance. The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes. Finally, results pretend to highlight the organisation of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.

@@

This research was partly funded by the European Research Council Advanced Grant EPNet (340828). XRC is funded by the Ramon y Cajal programme RYC2018-024050-I (Fondo Social Europeo – Agencia Estatal de Investigación).
We are grateful to Simon Carrignon, Juan Moros, Ignacio Morer and Víctor M. Lozano for their useful suggestions.  
Data have been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}. Maps were performed using QGIS software v.3.14 `Pi'. Datasets and source code are freely available under Open licenses from https://github.com/Mcotsar/Baetica\_stamps.
